---
title: "My Finances - summary report"
author: "Aiswarya Prasad"
date: "`r Sys.Date()`"
runtime: shiny
output:
#     Features matrix
#     Responsive	Dynamic TOC	Dark mode	Thumbnails / Lightbox	Code folding	Tabsets	Bad joke
#     html_docco	x			x	x	x
#     html_clean	x	x		x	x	x
#     readthedown	x	x			x	x
#     material				x	x	x
#         robobook	x	x		x	x	x
#     downcute	x	x	x	x	x	x
#     lockdown							x
# for more info see https://juba.github.io/rmdformats/
 rmdformats::robobook:
  #     fig_width : figures width, in inches
  #     fig_height : figures height, in inches
  #     fig_caption : toggle figure caption rendering
  #     highlight : syntax highlighting
  #     thumbnails : if TRUE, display content images as thumbnails
  #     lightbox : if TRUE, add lightbox effect to content images
  #     gallery : if TRUE, add navigation between images when displayed in lightbox
  #     use_bookdown : if TRUE, will use bookdown instead of rmarkdown for HTML rendering, thus providing section numbering and cross references.
  #     embed_fonts : if TRUE (default), use local files for fonts used in the template instead of links to Google Web fonts. This leads to bigger files but ensures that the fonts are available
  # additional aguments are passed to the base html_document RMarkdown template
  fig_caption : true
  lightbox: true
  code_folding: hide
  df_print: paged
  toc_float:
    collapsed: false
    smooth_scroll: true
---

```{r setup, include=F, echo=F}
# set the root directory up correctly
ROOT = getwd()
# load libraries
library(xlsx)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(magrittr)
library(lubridate)
library(ggthemes)
library(RColorBrewer)
library(shiny)
library(DT)
library(leaflet)
```

```{r function_and_defs, echo=F}
#########################################
# set colors and theme
#########################################
# maxcolors category colorblind
# BrBG            11      div       TRUE
# PiYG            11      div       TRUE
# PRGn            11      div       TRUE
# PuOr            11      div       TRUE
# RdBu            11      div       TRUE
# RdGy            11      div      FALSE
# RdYlBu          11      div       TRUE
# RdYlGn          11      div      FALSE
# Spectral        11      div      FALSE
# Accent           8     qual      FALSE
# Dark2            8     qual       TRUE
# Paired          12     qual       TRUE
# Pastel1          9     qual      FALSE
# Pastel2          8     qual      FALSE
# Set1             9     qual      FALSE
# Set2             8     qual       TRUE
# Set3            12     qual      FALSE
# Blues            9      seq       TRUE
# BuGn             9      seq       TRUE
# BuPu             9      seq       TRUE
# GnBu             9      seq       TRUE
# Greens           9      seq       TRUE
# Greys            9      seq       TRUE
# Oranges          9      seq       TRUE
# OrRd             9      seq       TRUE
# PuBu             9      seq       TRUE
# PuBuGn           9      seq       TRUE
# PuRd             9      seq       TRUE
# Purples          9      seq       TRUE
# RdPu             9      seq       TRUE
# Reds             9      seq       TRUE
# YlGn             9      seq       TRUE
# YlGnBu           9      seq       TRUE
# YlOrBr           9      seq       TRUE
# YlOrRd           9      seq       TRUE
#       maxcolors category colorblind
# Blues         9      seq       TRUE
# [1] 9
make_theme <- function(theme_name=theme_classic() ,max_colors=0, palettefill="Pastel1", palettecolor="Dark2",
                        setFill=TRUE, setCol=TRUE,
                        guide_nrow=2, guide_nrow_byrow=TRUE, leg_pos="top", leg_size=12,
                        x_angle=0 ,x_vj=0, x_hj=0, x_size=12,
                        y_angle=0 ,y_vj=0, y_hj=0, y_size=12){
  n_11 = c("BrBG", "PiYG", "PRGn", "PuOr", "RdBu", "RdGy", "RdYlBu", "RdYlGn", "Spectral")
  n_12 = c("Paired", "Set3")
  n_8 = c("Accent", "Dark2", "Pastel2", "Set2")
  if (palettefill %in% n_12) {
    n_f = 12
  } else {
    if (palettefill %in% n_11) {
      n_f = 11
    } else {
      if (palettefill %in% n_8) {
        n_f  = 8
      } else {
        n_f = 9
      }
    }
  }
  if (palettecolor %in% n_12) {
    n_c = 12
  } else {
    if (palettecolor %in% n_11) {
      n_c = 11
    } else {
      if (palettecolor %in% n_8) {
        n_c  = 8
      } else {
        n_c = 9
      }
    }
  }
  getFill = colorRampPalette(brewer.pal(n_f, palettefill))
  getColor = colorRampPalette(brewer.pal(n_c, palettecolor))
  theme_params <- theme(axis.text.x = element_text(angle = x_angle,
    vjust = x_vj, hjust=x_hj,
    size = x_size),
    axis.text.y = element_text(angle = y_angle,
      vjust = y_vj, hjust=y_hj,
      size = y_size),
      # axis.title.x = element_text(margin=margin(t=5)),
      # axis.title.y = element_text(margin=margin(r=10)),
      legend.position=leg_pos,
      legend.text = element_text(size=leg_size)
    )
  guide_params <- guides(fill = guide_legend(
                                  nrow=guide_nrow,
                                  byrow=guide_nrow_byrow
                                ),
                        col = guide_legend(
                                  nrow=guide_nrow,
                                  byrow=guide_nrow_byrow
                                )
                  )
  my_theme <- list(
                theme_name,
                theme_params,
                guide_params
              )

  if(setFill) {
    if (n_f < max_colors) {
      my_theme <- list(
                    my_theme,
                    scale_fill_manual(values = getFill(max_colors))
                  )

    } else {
      my_theme <- list(
                    my_theme,
                    scale_fill_brewer(palette=palettefill)
                  )
    }
  }
  if(setCol) {
    if (n_c < max_colors) {
      my_theme <- list(
                    my_theme,
                    scale_color_manual(values = getColor(max_colors))
                  )

    } else {
      my_theme <- list(
                    my_theme,
                    scale_color_brewer(palette=palettecolor)
                  )
    }
  }
  return(my_theme)
}

#########################################
# set next function
#########################################
```

```{css echo=F}
/* To make hoverable links. (does not have to be called hint) Usage: */
/* [Message to show on hover]{.hint} */
.hint {
  visibility: hidden;
}

.hint::before {
  visibility: visible;
  content: "Hint";
  color: blue;
}

.hint:hover {
  visibility: visible;
  font-weight: bold;
}

.hint:hover::before {
  display: none;
}
```

```{r read_data, echo=F}
# sheet1 is "All_Transactions"
all_df <- read.xlsx(paste0(ROOT, "/../Example-All_Transactions.xlsx"), 1, header=TRUE)
cols_to_choose <- c("Booking.date", "Notification.text", "Credit.in.CHF", "Debit.in.CHF", "Category", "Sub.category", "Notes")
all_df <- all_df[, cols_to_choose]
colnames(all_df) <- c("Date", "Description", "Credit", "Debit", "Category", "Subcategory", "Notes")
all_df <- all_df %>%
            replace_na(list(Credit = 0, Debit = 0))
df <- all_df
df <- filter(df, !is.na(Date))
df <- filter(df, !is.na(Description))

CategoriesToExclude <- c("Reimbursed", "Miscellaneous")

excluded_df <- df %>%
                filter(Category %in% CategoriesToExclude)
df <- df %>%
        filter(!Category %in% CategoriesToExclude)
all_df <- all_df %>%
        filter(!Category %in% CategoriesToExclude)
```

```{r prelim_analysis, echo=F}
salary_dates <- df %>%
                  filter(Category == "Salary") %>%
                    pull(Date)
num_salaries <- length(salary_dates)
min_date <- min(df$Date)
max_date <- max(df$Date)
last_year_start <- paste0(as.numeric(format(Sys.Date(), "%Y"))-1, "-01-01")
if (last_year_start < min_date) {
  last_year_start = min_date
}
date_order <- format(seq(floor_date(min_date, unit="month"), ceiling_date(max_date, unit="month"), by = "month"), "%h-%Y")
# account balance from all_df

```

# Transactions

This report includes transactions from **`r day(min_date)`, `r month(min_date, label=T, abbr=F)` `r year(min_date)`**  to **`r day(max_date)`, `r month(max_date, label=T, abbr=F)` `r year(max_date)`**.

For a specific date range use:

```{r set_date_range, echo=F}
dateRangeInput("date_range", "Date range:", start = min_date,
               end = max_date, min = min_date, max = max_date,
               format = "yyyy-mm-dd", startview = "month", weekstart = 0,
               language = "en", separator = " to ", width = "100%"
)
actionButton("Reset", "Reset", icon = NULL, width = NULL)
observe({
    input$Reset
    updateDateRangeInput(session, "date_range", "Date range:",,
      start = min_date,
      end = max_date
    )
})
df_reactive <- reactive({
                df %>%
                  filter(Date >= input$date_range[1] &
                    Date <= input$date_range[2])
})
```

## Total income and expense each month

```{r income_expense_plot, echo=F}
radioButtons("ShowLabelsIncomeExpense", "Show values on plot:",
             choices = c("Income" = "i", "Expense" = "e", "None" = "n"),
             inline = T,
             selected = "n"
)

renderPlot({
  df_plot <- df_reactive()
  df_plot <- df_plot %>%
                mutate(Date = paste0(format(Date, "%h"), "-", format(Date, "%Y"))) %>%
                  group_by(Date) %>%
                    summarise(income = sum(Credit), expense = sum(-Debit))

  p <- ggplot(data = pivot_longer(df_plot, 2:3, names_to = "Type", values_to = "Amount"),
              aes(x = factor(Date, levels=date_order), y = Amount, fill = Type)) +
            geom_bar(stat="identity",
                     color="grey",
                     position=position_dodge()) +
              ylab("Amount in CHF") +
                xlab("Month") +
                  make_theme(theme_few(), leg_pos="right", setFill=T,
                             x_angle=70, x_size=15, x_hj=1, x_vj=1,
                             y_size=15
                  )
  if (input$ShowLabelsIncomeExpense == "i") {
    p +
      geom_text(aes(label = ifelse(Type == "income", paste0(round(Amount, 0), "\nCHF"), ""),
                    y = Amount + 500,
                    lineheight = 0.75
              )
      ) +
      geom_text(aes(label = ifelse(Type == "income", paste0(round(Amount, 0), "\nCHF"), ""),
                    y = Amount + 500,
                    lineheight = 0.75
              )
      ) +
      geom_text(aes(label = ifelse(Type == "income", paste0(round(Amount, 0), "\nCHF"), ""),
                    y = Amount + 500,
                    lineheight = 0.75
              )
      )
  }
  else if (input$ShowLabelsIncomeExpense == "e") {
    p +
      geom_text(aes(label = ifelse(Type == "expense", paste0(round(Amount, 0), "\nCHF"), ""),
                    y = Amount + 800,
                    lineheight = 0.75
              )
      ) +
      geom_text(aes(label = ifelse(Type == "expense", paste0(round(Amount, 0), "\nCHF"), ""),
                    y = Amount + 800,
                    lineheight = 0.75
              )
      ) +
      geom_text(aes(label = ifelse(Type == "expense", paste0(round(Amount, 0), "\nCHF"), ""),
                    y = Amount + 800,
                    lineheight = 0.75
              )
      )
  }
  else if (input$ShowLabelsIncomeExpense == "n") {
    p
  }
}, height = 400, width = 800)

actionButton("This_year1", "This year", icon = NULL, width = NULL)
observe({
    input$This_year1
    updateDateRangeInput(session, "date_range", "Date range:",
      start = format(Sys.Date(), "%Y-01-01"),
      end = max_date
    )
})
actionButton("Last_year1", "Last year", icon = NULL, width = NULL)
observe({
    input$Last_year1
    updateDateRangeInput(session, "date_range", "Date range:",
      start = last_year_start,
      end = paste0(as.numeric(format(Sys.Date(), "%Y"))-1, "-12-31")
    )
})
actionButton("Show_all1", "Show all", icon = NULL, width = NULL)
observe({
    input$Show_all1
    updateDateRangeInput(session, "date_range", "Date range:",,
      start = min_date,
      end = max_date
    )
})
```

## Total savings each month

```{r savings_plot, echo=F}
radioButtons("ShowLabelsSavings", "Show values on plot:",
             choices = c("Yes" = "y", "No" = "n"),
             inline = T,
             selected = "n"
)

renderPlot({
  df_plot <- df_reactive()
  df_plot <- df_plot %>%
                mutate(Date = paste0(format(Date, "%h"), "-", format(Date, "%Y"))) %>%
                  group_by(Date) %>%
                    summarise(income = sum(Credit), expense = sum(-Debit)) %>%
                      summarise(Date, Savings = income - expense)

  p <- ggplot(data = df_plot,
              aes(x = factor(Date, levels=date_order),
                  y = Savings)) +
            geom_bar(stat="identity",
                     width = 0.8,
                     position=position_dodge(width = 0.5),
                     color="grey",
                     fill = brewer.pal(8, "Pastel2")[2]) +
              ylab("Amount in CHF") +
                xlab("Month") +
                  make_theme(theme_few(), leg_pos="right", setFill=F,
                             x_angle=70, x_size=15, x_hj=1, x_vj=1,
                             y_size=15
                  )
  if (input$ShowLabelsSavings == "y") {
    p +
      geom_text(aes(label=paste0(round(Savings, 0), "\nCHF"),
                  y = Savings + 300 * sign(Savings),
                  lineheight = 0.75
              )
      ) +
        geom_text(aes(label=paste0(round(Savings, 0), "\nCHF"),
                      y = Savings + 300 * sign(Savings),
                      lineheight = 0.75
                  )
        ) +
      geom_text(aes(label=paste0(round(Savings, 0), "\nCHF"),
                  y = Savings + 300 * sign(Savings),
                  lineheight = 0.75
              )
      )
  }
  else if (input$ShowLabelsSavings == "n") {
    p
  }
}, height = 400, width = 800)

actionButton("This_year2", "This year", icon = NULL, width = NULL)
observe({
    input$This_year2
    updateDateRangeInput(session, "date_range", "Date range:",
      start = format(Sys.Date(), "%Y-01-01"),
      end = max_date
    )
})
actionButton("Last_year2", "Last year", icon = NULL, width = NULL)
observe({
    input$Last_year2
    updateDateRangeInput(session, "date_range", "Date range:",
      start = last_year_start,
      end = paste0(as.numeric(format(Sys.Date(), "%Y"))-1, "-12-31")
    )
})
actionButton("Show_all2", "Show all", icon = NULL, width = NULL)
observe({
    input$Show_all2
    updateDateRangeInput(session, "date_range", "Date range:",,
      start = min_date,
      end = max_date
    )
})
div(renderDataTable({
    df_filtered <- df_reactive()
    df_filtered <- df_filtered %>%
                    filter(Date >= input$date_range[1] &
                      Date <= input$date_range[2]) %>%
                      select(!Notes)
    df_savings <- df_filtered %>%
                    mutate(Date = paste0(format(Date, "%h"), "-", format(Date, "%Y"))) %>%
                      group_by(Date) %>%
                        summarise(income = sum(Credit), expense = sum(-Debit)) %>%
                          summarise(Date, Savings = income - expense)
    avg_savings <- round(mean(df_savings$Savings), 2)
    df_income_expense <- df_filtered %>%
                          mutate(Date = paste0(format(Date, "%h"), "-", format(Date, "%Y"))) %>%
                            group_by(Date) %>%
                              summarise(income = sum(Credit), expense = sum(-Debit))
    avg_income <- round(mean(df_income_expense$income), 2)
    avg_expense <- round(mean(df_income_expense$expense), 2)

    df_avg <- data.frame("Average" = c("Income", "Expense", "Savings"),
                         "Amount" = c(avg_income, avg_expense, avg_savings))
    df_avg
}, rownames = F, colnames = c("Average", "Amount in CHF"), options = list(dom = 't')), style = "font-size:150%")
actionButton("This_year_table", "This year", icon = NULL, width = NULL)
observe({
    input$This_year_table
    updateDateRangeInput(session, "date_range", "Date range:",
      start = format(Sys.Date(), "%Y-01-01"),
      end = max_date
    )
})
actionButton("Last_year_table", "Last year", icon = NULL, width = NULL)
observe({
    input$Last_year_table
    updateDateRangeInput(session, "date_range", "Date range:",
      start = last_year_start,
      end = paste0(as.numeric(format(Sys.Date(), "%Y"))-1, "-12-31")
    )
})
actionButton("Show_all_table", "Show all", icon = NULL, width = NULL)
observe({
    input$Show_all_table
    updateDateRangeInput(session, "date_range", "Date range:",,
      start = min_date,
      end = max_date
    )
})
```

<!-- ## Expenditure by categories -->


# Transaction table

```{r show_all_data, echo=F}
renderDataTable({
  df_reactive() %>%
    filter(Date >= input$date_range[1] &
      Date <= input$date_range[2]) %>%
      select(!Notes) %>%
        pivot_longer(cols = 3:4, names_to = "type", values_to = "Amount") %>%
          select(!type) %>%
            filter(Amount != 0)
}, filter = "top", rownames = F)
actionButton("This_year_df", "This year", icon = NULL, width = NULL)
observe({
    input$This_year_df
    updateDateRangeInput(session, "date_range", "Date range:",
      start = format(Sys.Date(), "%Y-01-01"),
      end = max_date
    )
})
actionButton("Last_year_df", "Last year", icon = NULL, width = NULL)
observe({
    input$Last_year_df
    updateDateRangeInput(session, "date_range", "Date range:",
      start = last_year_start,
      end = paste0(as.numeric(format(Sys.Date(), "%Y"))-1, "-12-31")
    )
})
actionButton("Show_all_df", "Show all", icon = NULL, width = NULL)
observe({
    input$Show_all_df
    updateDateRangeInput(session, "date_range", "Date range:",,
      start = min_date,
      end = max_date
    )
})

```


# Disclaimers


The transactions are excluded from reports.

`r renderDataTable({excluded_df %>% select(!Notes) %>% pivot_longer(cols = 3:4, names_to = "type", values_to = "Amount") %>% select(!type) %>% filter(Amount != 0)}, filter = "top", rownames = F)`

To include these transactions, find and change the line `CategoriesToExclude <- c("Reimbursed", "Miscellaneous")` in the R markdown document.

In the future, if particular Miscellaneous transactions are to be excluded, they need to be mentioned seperately (i.e. the code excluding transactions should be modified to consider them). Any transactions flagged as one of the categories to be excluded will be excluded.

# Extra features that would be nice to implement

* For now, more plots!

* Interactive plots?

* Projected expenses based on a monthly expense that might be made
